---
title: "Nuggets vs. Jazz Lineups"
author: "kyle wurtz"
date: "2019.06.28"
output: 
  html_notebook:
    toc: true
    toc_float: true
    theme: yeti
    code_folding: hide
---

## Overview

## Setup

### Load Packages
```{r load packages}
require(tidyverse)
require(magrittr)
require(here)
```

### Inputs

```{r}
input_data = "/Users/kylewurtz/Dropbox/R/NBA Analysis/NBA PBP Data/2018-2019_NBA_PbP_Logs/[10-16-2018]-[06-13-2019]-combined-stats.csv"
```

## Data Prep

```{r, results = "hide"}
df = read_csv(input_data)
```

```{r}
head(df)
```

Filter to just the games where the Nuggets played the Jazz:

```{r}
nuggets_game_ids = df %>% 
  filter(team == "DEN") %>% 
  select(game_id) %>% 
  distinct() %>% 
  mutate(game_id = gsub('"', "", gsub('="', "", game_id)))
jazz_game_ids = df %>% 
  filter(team == "UTA") %>% 
  select(game_id) %>% 
  distinct() %>% 
  mutate(game_id = gsub('"', "", gsub('="', "", game_id)))
nugs_vs_jazz_games = df %>% 
  mutate(game_id = gsub('"', "", gsub('="', "", game_id))) %>% 
  filter((game_id %in% nuggets_game_ids$game_id) & (game_id %in% jazz_game_ids$game_id))
```

Figure out if the Nuggets were home or away (How the f does BigDataBall not include this info in their data dump? Seriously...that's so pathetic...).

```{r}
home_or_away_ind = nugs_vs_jazz_games %>% 
  select(game_id, a1:h5) %>% 
  gather(key, value, -game_id) %>% 
  filter(value == "Jamal Murray") %>% 
  mutate(home_or_away = str_sub(key, 1, 1)) %>% 
  select(game_id, home_or_away) %>% 
  distinct()
home_or_away_ind
```

Join indicator onto data set:

```{r}
nugs_vs_jazz_games %<>%
  left_join(
    home_or_away_ind %>% 
      rename(nuggets_home_or_away_ind = home_or_away)
  )
```

## Analysis

### Millsap and Jokic

How did the Nuggets perform against the Jazz while both Millsap and Jokic were on the court?

Start by filtering to possessions where both were on the court:

```{r}
nugs_vs_jazz_games %<>% 
  mutate(
    jokic_on_court_ind = ifelse(
      nuggets_home_or_away_ind == "h",
      # home team
      ifelse(
        (h1 == "Nikola Jokic" | h2 == "Nikola Jokic" | h3 == "Nikola Jokic" |
           h4 == "Nikola Jokic" | h5 == "Nikola Jokic"),
        1,
        0
      ),
      # away team
      ifelse(
        (a1 == "Nikola Jokic" | a2 == "Nikola Jokic" | a3 == "Nikola Jokic" |
           a4 == "Nikola Jokic" | a5 == "Nikola Jokic"),
        1,
        0
      )
    ),
    millsap_on_court_ind = ifelse(
      nuggets_home_or_away_ind == "h",
      # home team
      ifelse(
        (h1 == "Paul Millsap" | h2 == "Paul Millsap" | h3 == "Paul Millsap" |
           h4 == "Paul Millsap" | h5 == "Paul Millsap"),
        1,
        0
      ),
      # away team
      ifelse(
        (a1 == "Paul Millsap" | a2 == "Paul Millsap" | a3 == "Paul Millsap" |
           a4 == "Paul Millsap" | a5 == "Paul Millsap"),
        1,
        0
      )
    ),
    gobert_on_court_ind = ifelse(
      nuggets_home_or_away_ind == "a",
      # jazz are home team
      ifelse(
        (h1 == "Rudy Gobert" | h2 == "Rudy Gobert" | h3 == "Rudy Gobert" |
           h4 == "Rudy Gobert" | h5 == "Rudy Gobert"),
        1,
        0
      ),
      # jazz are away team
      ifelse(
        (a1 == "Rudy Gobert" | a2 == "Rudy Gobert" | a3 == "Rudy Gobert" |
           a4 == "Rudy Gobert" | a5 == "Rudy Gobert"),
        1,
        0
      )
    ),
    nugs_score = ifelse(
      nuggets_home_or_away_ind == "h",
      home_score,
      away_score
    ),
    jazz_score = ifelse(
      nuggets_home_or_away_ind == "h",
      away_score,
      home_score
    ),
    nugs_score_advantage = nugs_score - jazz_score
  )
mj_vs_jazz = nugs_vs_jazz_games %>% 
  filter(jokic_on_court_ind == 1, millsap_on_court_ind == 1)
mj_vs_jazz %>% 
  count(game_id)
```

```{r}
mj_vs_jazz %>% 
  group_by(game_id, team) %>% 
  summarize(
    points = sum(points, na.rm = T)
  ) %>% 
  filter(!is.na(team)) %>% 
  spread(team, points) %>% 
  mutate(nugs_point_differential = DEN - UTA)
```

```{r}
mj_vs_jazz %>% 
  group_by(team) %>% 
  summarize(
    points = sum(points, na.rm = T)
  ) %>% 
  filter(!is.na(team)) %>% 
  spread(team, points) %>% 
  mutate(nugs_point_differential = DEN - UTA)
```


What about with Gobert on the court?

```{r}
mj_vs_jazz %>% 
  filter(gobert_on_court_ind == 1) %>% 
  group_by(game_id, team) %>% 
  summarize(
    points = sum(points, na.rm = T)
  ) %>% 
  filter(!is.na(team)) %>% 
  spread(team, points) %>% 
  mutate(nugs_point_differential = DEN - UTA)
```

Gobert off the court:

```{r}
mj_vs_jazz %>% 
  filter(gobert_on_court_ind == 0) %>% 
  group_by(game_id, team) %>% 
  summarize(
    points = sum(points, na.rm = T)
  ) %>% 
  filter(!is.na(team)) %>% 
  spread(team, points) %>% 
  mutate(nugs_point_differential = DEN - UTA)
```

Just Jokic and Gobert:

```{r}
nugs_vs_jazz_games %>% 
  filter(jokic_on_court_ind == 1, gobert_on_court_ind == 1) %>% 
  group_by(team) %>% 
  summarize(
    points = sum(points, na.rm = T)
  ) %>% 
  filter(!is.na(team)) %>% 
  spread(team, points) %>% 
  mutate(nugs_point_differential = DEN - UTA)
# nugs_vs_jazz_games %>% 
#   filter(jokic_on_court_ind == 1, gobert_on_court_ind == 1) %>% 
#   count()
```

Jokic without Gobert:

```{r}
nugs_vs_jazz_games %>% 
  filter(jokic_on_court_ind == 1, gobert_on_court_ind == 0) %>% 
  group_by(team) %>% 
  summarize(
    points = sum(points, na.rm = T)
  ) %>% 
  filter(!is.na(team)) %>% 
  spread(team, points) %>% 
  mutate(nugs_point_differential = DEN - UTA)
# nugs_vs_jazz_games %>% 
#   filter(jokic_on_court_ind == 1, gobert_on_court_ind == 0) %>% 
#   count()
```

```{r}
nugs_vs_jazz_games %>% 
  filter(jokic_on_court_ind == 1) %>% 
  group_by(team) %>% 
  summarize(
    points = sum(points, na.rm = T)
  ) %>% 
  filter(!is.na(team)) %>% 
  spread(team, points) %>% 
  mutate(nugs_point_differential = DEN - UTA)
```


```{r}
nugs_vs_jazz_games %>% 
  group_by(game_id) %>% 
  slice(n()) %>% 
  select(home_score, away_score, nuggets_home_or_away_ind)
```





